Visualising metagenome differences

### 2023-11-20 analysis/metagenome.Rmd
Add reformatted CAT for easier plotting

>>>>

```{r}
metadata <- read_tsv('./data/metagenome/samples.tsv')
metadata <- metadata |> rename(Sample_species = Species,
                               Sample_replicate = Replicate) |> 
  mutate(Tissue = str_to_title(Tissue))
```

# Taxonomies

We have two taxonomies: one based on CAT, one based on GTDB. GTDB is less complete but more accurate. Let's see what they look like.

# GTDB
```{r}
tax <- read_tsv('./data/metagenome/all_gtdb.summary.tsv.gz')
```

```{r}
clean_tax <- tax |> 
  separate(user_genome, into = c('sample', 'MAG'), sep='\\.') |> 
  separate(classification, into = c('domain', 'phylum', 
                                    'class', 'order', 
                                    'family', 'genus', 
                                    'species'), sep=';') |> 
  mutate(sample = str_replace_all(sample, 'MEGAHIT-', ''), 
         sample = str_replace_all(sample, '_L2', ''))

clean_tax <- clean_tax |> left_join(metadata, by = c('sample'='Sample'))
```

Next, we keep only Posidonia australis and Amphibolis antarctica samples

```{r}
amph_pos_tax <- clean_tax |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
```

## GTDB by order

```{r}
amph_pos_tax |> 
  group_by(sample, Tissue) |> 
  mutate(order = str_replace_all(order, 'o__','')) |> 
  count(order) |> 
  filter ( n > 1) |> 
  filter(order != '' ) |> 
  ggplot(aes(x = sample, y = n, fill=order)) +
  geom_col() + 
  theme_minimal() + 
  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  facet_wrap(~Tissue) + coord_flip() + 
  ylab('Number of contigs') + xlab('Sample')
```

## GTDB by family

```{r}
amph_pos_tax |> 
  group_by(sample, Tissue) |> 
  mutate(family = str_replace_all(family, 'f__','')) |> 
  count(family) |> 
  filter ( n > 1) |> 
  filter(family != '' ) |> 
  ggplot(aes(x = sample, y = n, fill=family)) +
  geom_col() + 
  theme_minimal() + 
  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  facet_wrap(~Tissue) + coord_flip() + 
  ylab('Number of contigs') + xlab('Sample')
```
Let's do the same using CAT

# CAT

```{r}
tax_cat <- read_tsv('./data/metagenome/all_cat.summary.tsv.gz', guess_max = 100000)
```

```{r}
clean_tax_cat <- tax_cat |> 
  separate(`# bin`, into = c('sample', 'MAG'), sep='\\.') |> 
  separate(full_lineage_names, into = c('root', 'root2', 
                                    'superkingdom', 'clade', 
                                    'phylum', 'subphylum', 
                                    'class'), sep='\t') |> 
  mutate(sample = str_replace_all(sample, 'MEGAHIT-', ''), 
         sample = str_replace_all(sample, '_L2', ''))

clean_tax <- clean_tax |> left_join(metadata, by = c('sample'='Sample'))
```

Next, we keep only Posidonia australis and Amphibolis antarctica samples

```{r}
amph_pos_tax <- clean_tax |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
```
